---
title: "Scrape get grand slam titles by age"
author: "Duc-Quang Nguyen"
date: "30 Jan 2017"
output: html_document
---

# Data

* [Most Tennis Grand Slam Titles Winners ](http://www.totalsportek.com/tennis/grand-slam-titles-winners-mens-women/)
* [Most Grand Slam singles titles (5 or more) in the open era](https://en.wikipedia.org/wiki/List_of_Grand_Slam_men%27s_singles_champions#Most_Grand_Slam_singles_titles_.285_or_more.29)
# Scrape

## ATP world tour

* [ATP ranking singles](http://www.atpworldtour.com/en/rankings/singles)
* [all former n°1](http://www.atpworldtour.com/en/rankings/former-no-1s)

## Related
 * [NZZ Der Weg zu Titel Nummer 18 in einer Grafik](https://www.nzz.ch/sport/tennis/roger-federer-der-weg-zu-titel-nummer-18-in-einer-grafik-ld.142428?reduced=true)
 * [Novak Djokovic Is Falling Behind In His Quest To Break Roger Federer’s Record](https://fivethirtyeight.com/features/novak-djokovic-is-falling-behind-in-his-quest-to-break-roger-federers-record/)
 * [Quora](https://www.quora.com/Can-Novak-Djokovic-surpass-Federer-and-Nadals-Grand-Slam-titles)
 
![graph](https://qph.ec.quoracdn.net/main-qimg-770d9868e7290610d5feea184e74633a-c) 

## rvest

 * [rvest tutorial: scraping the web using R](http://stat4701.github.io/edav/2015/04/02/rvest_tutorial/)


```{r setup, include=FALSE}
getData <- F
translation.file <- "input/Grand slam title by age - Sheet1.csv"

library(tidyverse)
library(rvest)
library(magrittr)


grandSlams <- c('Australian Open', 'Wimbledon', 'Roland Garros', 'US Open')
names(grandSlams) <- c('01-29', '07-16', '06-10', '09-10')

# all players with at least 3 Grand Slam singles titles during the open era

players <- c(
  # past players from 
  # 1. https://en.wikipedia.org/wiki/List_of_Grand_Slam_men%27s_singles_champions#Most_singles_titles_by_Grand_Slam_tournament
  # 2. For players with 3 or more https://www.quora.com/Can-Novak-Djokovic-surpass-Federer-and-Nadals-Grand-Slam-titles
  #  Kuerten, Ash, Villas, Courier
  'http://www.atpworldtour.com//en/players/pete-sampras/s402/overview',
  'http://www.atpworldtour.com/en/players/bjorn-borg/b058/overview',
  'http://www.atpworldtour.com/en/players/jimmy-connors/c044/overview',
  'http://www.atpworldtour.com/en/players/ivan-lendl/l018/overview',
  'http://www.atpworldtour.com/en/players/andre-agassi/a092/overview',
  'http://www.atpworldtour.com/en/players/john-mcenroe/m047/overview',
  'http://www.atpworldtour.com/en/players/mats-wilander/w023/overview',
  'http://www.atpworldtour.com/en/players/stefan-edberg/e004/overview',
  'http://www.atpworldtour.com/en/players/boris-becker/b028/overview',
  'http://www.atpworldtour.com/en/players/rod-laver/l058/overview',
  'http://www.atpworldtour.com/en/players/john-newcombe/n044/overview',
  'http://www.atpworldtour.com/en/players/gustavo-kuerten/k293/overview',
  'http://www.atpworldtour.com/en/players/arthur-ashe/a063/overview',
  'http://www.atpworldtour.com/en/players/guillermo-vilas/v028/overview',
  'http://www.atpworldtour.com/en/players/jim-courier/c243/overview',
  # add 
  'http://www.atpworldtour.com/en/players/roger-federer/f324/overview',
  'http://www.atpworldtour.com/en/players/rafael-nadal/n409/overview',
  'http://www.atpworldtour.com/en/players/novak-djokovic/d643/overview',
  'http://www.atpworldtour.com/en/players/stan-wawrinka/w367/overview',
  'http://www.atpworldtour.com/en/players/andy-murray/mc10/overview'
)

players <- tibble(
  name =  gsub("^.*players\\/.*\\-(.*)\\/.*\\/overview", "\\1", players),
  url = players, 
  color = '#beb298',
  until = NA)

#col.player <- c('#ab3d3f', '#666666', '#333366', '#d4c3aa', '#808080', '#999999')
#names(col.player) <- names(players)

```

## Download all titles

** Get Data ** 

```{r get path to titles}
if(getData) {
  
  ## 1. Get list of titles
  links <- gsub("overview", "titles-and-finals", players$url)
  names(links) <- players$name
  
  link <- links[4]
  titles <- do.call(rbind, lapply(links, function(link) {
    cat("\nGet list of titles", link)
    rtable <- read_html(link) %>% 
      html_nodes(".mega-table") %>% html_table()
    rtable <- rtable[[1]]
    
    years <- rtable[[1]]
    gs <- sapply(grandSlams, function(slam) grep(slam, rtable[[3]]))
    gs <- sapply(gs, function(ii) if(length(ii) > 0) years[ii])
    # remove NULL elements
    gs <- gs[!sapply(gs, is.null)] 
    
    gs <- do.call(rbind, lapply(names(gs), function(slam) {
      data.frame(
        gs = as.character(grandSlams[which(names(grandSlams) == slam)]), 
        date = as.Date(paste0(gs[[slam]], "-", names(grandSlams)[which(names(grandSlams) == slam)]))
      )
    }))
    gs$player <- names(links)[which(links == link)]
    gs
  }))
  
  rownames(titles) <- NULL
 
  ## 2. Get retirement date 
  links <- gsub("overview", "rankings-history", players$url)
  names(links) <- players$name
  
  playingUntil <- sapply(links, function(link) {
    cat("\nGet retirement age", link) 
    rtable <- read_html(link) %>% 
    html_nodes(".mega-table") %>% html_table()
  
    rtable <- rtable[[1]] %>% select (-Doubles)
    rtable$Date <- as.Date(rtable$Date, format = "%Y.%m.%d")
    rtable$Date[1]
  })
  playingUntil <- as.Date(playingUntil, origin = "1970-01-01") %>% 
    enframe() %>% rename(playingUntil = value)
  
  
  ## 3. Get player DOB
  links <- players$url
  names(links) <- players$name
  dob <- sapply(links, function(url) {
    cat("\nGet date of birth", url)
    dob <- read_html(url) %>% 
      html_nodes(".table-birthday")
    gsub(".*(\\d{4}\\.\\d{2}\\.\\d{2}).*", "\\1", dob)
  }) 
  
  dob <- structure(as.Date(dob, format = "%Y.%m.%d"), names = names(dob))
  dob %<>% enframe() %>% 
    rename(dob = value, player = name)
    
  ## Combine data!
  
  # add 18 years old observation
  titles <- rbind(titles, data.frame(
    gs = NA, 
    date = dob$dob + (365.25 * 16),
    player = dob$player
  ))
  # add end (retirement age or current time)
  titles <- rbind(titles, data.frame(
    gs = NA, 
    date = playingUntil$playingUntil,
    player = playingUntil$name
  ))
    
  # add date of birth
  titles %<>% left_join(dob)
  
  ## r wrangle
  titles$age <- as.numeric((titles$date - titles$dob) / 365.25)
  titles$agef <- floor(titles$age)
  titles %<>% arrange(date)
  
  titles$title <- ifelse(is.na(titles$gs), 0, 1)
  
  # CumulatQive sum of grand slam titles
  titles %<>% 
    group_by(player) %>% 
    mutate(count = cumsum(title)) %>%
    ungroup()
  
  write.csv(titles, file = "input/titlesByAgeOpenEra.csv", row.names = F)  
} else {
  titles <- read_csv("input/titlesByAgeOpenEra.csv")
}

data <- titles %>% select(-dob)
data$colour <- players[match(data$player, players$name), "color"] %>% unlist(use.names = F)
```

```{r dev}
 ggplot(data = data) + 
    geom_step(aes(age, count, group = player, colour = colour), size = 1, alpha = 1)

```

## Following part is not working!

```{r plot}
library(ggplot2)
library(scales)
library(swiTheme)

txt <- loadTranslation(translation.file)

# firstup <- function(x) {
#    substr(x, 1, 1) <- toupper(substr(x, 1, 1))
# x
# }

lang <- 'EN'

myTheme <- function(base_family, title_family, base_size = 19) {
  swi_theme(base_family = base_family, subtitle = base_family, title_family = title_family, base_size = base_size) + 
    theme(
      panel.grid.major.y=element_line(color="#b3b3b3", linetype="dashed", size=0.2),
      panel.background = element_rect(fill = "#f2f2f2", color = NA),
      plot.background = element_rect(fill = "#f2f2f2", color = NA),
      strip.text = element_text(size = 12, hjust = 0, family = base_family, colour = "#595959", face = "bold"),
      legend.position = c(.83, .4),
      plot.subtitle=element_text(family = base_family, color = '#4A4A4A', size = base_size * 0.9),
      plot.caption=element_text(size = base_size * 0.6, margin=margin(t=13), color = '#737373', hjust = 1),
      plot.title=element_text(family = title_family, size = base_size * ifelse(lang == 'DE', 0.9, 1.1), margin=margin(b=13)),
      plot.margin=unit(c(0.6, 0.45, 0.05, 0.45), "cm")
    ) 
}

for(lang in colnames(txt)) {
  cat("\n", lang)
  ddd <- data
  
  # get translations
  ddd$player <- txt[ddd$player, lang]
  ddd$gs <- txt[ddd$gs, lang]
  
  gp <- ggplot(data = ddd) + 
    geom_line(aes(age, count, group = player, colour = colour), size = 1, alpha = 1) + 
    myTheme(base_family = txt['base.font', lang], title_family = txt['title.font', lang],
            base_size = 19 * as.numeric(txt['font.factor', lang])) +
    geom_point(
      data = ddd %>% filter(!is.na(gs)), 
      aes(age, count, group = gs, shape = gs, fill = gs), size = 3.2, 
      alpha = 0.77, colour = "white") +
    scale_colour_identity() + 
    scale_x_continuous(name = txt['x.lab', lang], expand = c(0.01,0.08)) +
    scale_y_continuous(name = txt['y.lab', lang], breaks = scales::pretty_breaks(n = 10), limits = c(0, 19), expand = c(0,0)) +
    scale_shape_manual(values = 21:24, guide = guide_legend(title = NULL))  + 
    scale_fill_manual(values = c('#3a9736', '#663333', '#333366', '#996699'), 
                      guide = guide_legend(title = NULL)) +
    geom_text(data = ddd %>% filter(is.na(gs), age > 20), 
              aes(age, count, label = player, colour = colour), 
              nudge_y = 0.29, hjust = 0.9, family = txt['title.font', lang], 
              size = 3.7 * as.numeric(txt['font.factor', lang])) +
    labs(
      subtitle = txt['descr', lang],
      caption = paste0(txt['source', lang], ": ", txt['source.name', lang], " | @duc_qn | swissinfo.ch "),
      title =  txt['main.title', lang])
  
  png(filename = paste0("output/graphic_openEra3orMoreGS_", lang, ".png") , width = 640 * 2, height = (640 * 2)*1.25, res = 170)
  print(gp)
  dev.off() 
}



```